\begin{thebibliography}{15}
\providecommand{\natexlab}[1]{#1}
\providecommand{\url}[1]{\texttt{#1}}
\expandafter\ifx\csname urlstyle\endcsname\relax
  \providecommand{\doi}[1]{doi: #1}\else
  \providecommand{\doi}{doi: \begingroup \urlstyle{rm}\Url}\fi

\bibitem[Albeverio and Hoegh-Krohn(1978)]{albeverio:1978}
S.~Albeverio and R.~Hoegh-Krohn.
\newblock Frobenius theory for positive maps on von {N}eumann algebras.
\newblock \emph{Communications in Mathematical Physics}, 64:\penalty0 83--94,
  1978.

\bibitem[Dunford and Schwartz(1958)]{dunfordschwartz:1958}
N.~Dunford and J.~Schwartz.
\newblock \emph{Linear Operators, {P}art {I}: {G}eneral {T}heory}.
\newblock Wiley, New York, 1958.

\bibitem[Evans(1977)]{evans:1977}
D.~Evans.
\newblock Irreducible quantum dynamical semigroups.
\newblock \emph{Communications in Mathematical Physics}, 54:\penalty0 293--297,
  1977.

\bibitem[Greiner(1981)]{greiner:1981}
G.~Greiner.
\newblock Zur {P}erron-{F}robenius {T}heorie stark stetiger {H}albgruppen.
\newblock \emph{Mathematische Zeitschrift}, 177:\penalty0 401--423, 1981.

\bibitem[Greiner(1982)]{greiner:1982}
G.~Greiner.
\newblock Spektrum und {A}symptotik stark stetiger {H}albgruppen positiver
  {O}peratoren.
\newblock \emph{Sitzungsberichte der Heidelberger Akademie der Wissenschaften,
  Mathematisch-Naturwissenschaftliche Klasse}, pages 55--80, 1982.

\bibitem[Greiner et~al.(1981)Greiner, Voigt, and Wolff]{greinervoigtwolff:1981}
G.~Greiner, J.~Voigt, and M.~Wolff.
\newblock On the spectral bound of the generator of semigroups of positive
  operators.
\newblock \emph{Journal of Operator Theory}, pages 245--256, 1981.

\bibitem[Groh(1981)]{groh:1981}
U.~Groh.
\newblock The peripheral point spectrum of {S}chwarz operators on
  {$C^*$}-algebras.
\newblock \emph{Mathematische Zeitschrift}, 176:\penalty0 311--318, 1981.

\bibitem[Groh(1982)]{groh:1982a}
U.~Groh.
\newblock Some observations on the spectra of positive operators on finite
  dimensional {$C^*$}-algebras.
\newblock \emph{Linear Algebra and its Applications}, 42:\penalty0 213--222,
  1982.

\bibitem[Groh(1984)]{groh:1984b}
U.~Groh.
\newblock Uniform ergodic theorems for identity preserving {S}chwarz maps on
  {W*}-algebras.
\newblock \emph{Journal of Operator Theory}, 11:\penalty0 395--404, 1984.

\bibitem[Hewitt and Ross(1963)]{hewittross:1963}
E.~Hewitt and K.~A. Ross.
\newblock \emph{Abstract Harmonic Analysis {I}}.
\newblock Springer, Berlin-Heidelberg-New York, 1963.

\bibitem[Jacobs(1972)]{jacobs:1972}
K.~Jacobs.
\newblock Gleichverteilung mod 1.
\newblock In \emph{Selecta Mathematica}, volume~IV, pages 57--93. Springer,
  Berlin-Heidelberg-New York, 1972.

\bibitem[Lin(1974)]{lin:1974}
M.~Lin.
\newblock On the uniform ergodic theorem {II}.
\newblock \emph{Proc. Amer. Math. Soc.}, 46:\penalty0 217--225, 1974.

\bibitem[Pedersen(1979)]{pedersen:1979}
G.~K. Pedersen.
\newblock \emph{{C*}-Algebras and their Automorphism Groups}.
\newblock Academic Press, London, New York, San Francisco, 1979.

\bibitem[Schaefer(1974)]{schaefer:1974}
H.~H. Schaefer.
\newblock \emph{Banach Lattices and Positive Operators}.
\newblock Springer, New York-Heidelberg-Berlin, 1974.

\bibitem[Takesaki(1979)]{takesaki:1979}
M.~Takesaki.
\newblock \emph{Theory of Operator Algebras {I}}.
\newblock Springer, New York-Heidelberg-Berlin, 1979.

\end{thebibliography}
