\bibliographystyle{spmpsci}
\begin{thebibliography}{11}
\providecommand{\natexlab}[1]{#1}
\providecommand{\url}[1]{\texttt{#1}}
\expandafter\ifx\csname urlstyle\endcsname\relax
  \providecommand{\doi}[1]{doi: #1}\else
  \providecommand{\doi}{doi: \begingroup \urlstyle{rm}\Url}\fi

\bibitem[Davies(1980)]{davies:1980}
E.~Davies.
\newblock \emph{One-parameter Semigroups}.
\newblock Academic Press, London-New York-San Francisco, 1980.

\bibitem[Engel and Nagel(2000)]{engelnagel:2000}
K.-J. Engel and R.~Nagel.
\newblock \emph{One-Paramter Semigroups for Linear Evolution Equations}.
\newblock Number 194 in Graduate Textes in Mathematics. Springer, Berlin AND
  Heidelberg AND New York, 2000.

\bibitem[Goldstein(1985{\natexlab{a}})]{goldstein:1985a}
J.~Goldstein.
\newblock \emph{Semigroups of Operators and Applications}.
\newblock Oxford University Press, 1985{\natexlab{a}}.

\bibitem[Goldstein(1985{\natexlab{b}})]{goldstein:1985b}
J.~Goldstein.
\newblock A (more-or-less) complete bibliography of semigroups of operators
  through 1984.
\newblock Preprints and lecture notes in mathematics, Tulane University,
  1985{\natexlab{b}}.

\bibitem[Hille and Phillips(1957)]{hillephillips:1957}
E.~Hille and R.~S. Phillips.
\newblock \emph{Functional Analysis and Semigroups}, volume~31 of
  \emph{Colloquium Publications}.
\newblock American Mathematical Society, Providence, R.I., 1957.

\bibitem[Kato(1966)]{kato:1966}
T.~Kato.
\newblock \emph{Perturbation Theory for Linear Operators}.
\newblock Springer, 1966.
\newblock 2nd printing: Berlin-Heidelberg-New York: Springer 1976.

\bibitem[Krein and Khazan(1985)]{kreinkhazan:1985}
S.~G. Krein and M.~I. Khazan.
\newblock Differential equations in a banach space.
\newblock \emph{J. Soviet Math.}, 30\penalty0 (3):\penalty0 2154--2239, 1985.
\newblock \doi{10.1007/BF02105398}.
\newblock URL \url{https://doi.org/10.1007/BF02105398}.

\bibitem[Pazy(1983)]{pazy:1983}
A.~Pazy.
\newblock \emph{Semigroups of Linear Operators and Applications to Partial
  Differential Equations}.
\newblock Springer, Berlin-Heidelberg-New York-Tokyo, 1983.

\bibitem[Phillips(1954)]{phillips:1954}
R.~S. Phillips.
\newblock A note on the abstract {Cauchy} problem.
\newblock \emph{Proc. Nat. Acad. Sci. U.S.A.}, 40:\penalty0 244--248, 1954.

\bibitem[Reed and Simon(1975)]{reedsimon:1975}
M.~Reed and B.~Simon.
\newblock \emph{Methods of Modern Mathematical Physics {II}. {Fourier}
  Analysis, Self-Adjointness}.
\newblock Academic Press, New York, 1975.

\bibitem[Schaefer(1974)]{schaefer:1974}
H.~H. Schaefer.
\newblock \emph{Banach Lattices and Positive Operators}.
\newblock Springer, New York-Heidelberg-Berlin, 1974.

\end{thebibliography}
