%% --
%% -- Updated Notes
%% -- Stand 2025-10-11
%% --
\section{Updated Notes D-II}
%% -- 
\begin{enumerate}
\item 
As we have seen in Chapter A-II, Section 3, strongly continuous semigroups on commutative \WA-algebras, that is, on $L^{\infty}$, are already norm-continuous. 
The proof depends heavily on the Grothendieck property (GP) and the Dunford-Pettis property (DPP) of these Banach spaces. 
It was shown by \mycite{zbMATH00537321} that every \WA-algebra has (GP) (see also \mycite{zbMATH00125315} for an alternative proof).
On the other hand, for a \WA-algebra $M$ the following are equivalent.
%% --
\begin{enumerate}[\upshape(a)]
    \item The \WA-algebra $M$ is of Type I finite.
    \item The \WA-algebra $M$ has (DPP).
    \item The predual $M_{*}$ has (DPP).
\end{enumerate}
%% --
(see \mycite{zbMATH00125315} and \mycite{zbMATH00097659}).
Thus, as a consequence of A-II Theorem 3.5, strongly continuous semigroups on Type I finite \WA-algebras have a bounded generator. 

\item
\mycite{zbMATH01495754} has shown that strong and norm convergence for sequences of completely positive maps are equivalent on \WA-algebras (even on the larger class of \AW-algebras).
Therefore a strongly continuous $C_{0}$-semigroup of completely positive operators on a \WA-algebra is norm continuous.

\item 
In \mycite{zbMATH00097715} it is shown that the equivalence of strong and norm convergence of semigroups on $C(K)$ implies the \CA-algebra $C(K)$ is a Grothendieck space.
This is also true for a noncommutative \CA-algebra $M$.
To prove this, one uses the fact that a \CA-algebra is a Grothendieck space if and only if $c_{0}$ is not a complemented subspace (see \mycite[Prop. 3.1.13 \& 4.2.1]{zbMATH07458830}).
Indeed, suppose that every strongly convergent $C_{0}$-semigroup is uniformly continuous on $M$.
Then $c_{0}$ can't be a complemented subspace of $\BA$.
For if $\BA = c_{0} \oplus F$ for a closed subspace $F$ of $M$, then take a strongly convergent but not uniformly convergent semigroup ${T(t)}$ on $c_{0}$ and let $S(t) = T(t) \oplus Id_{F}$ on $M$. 
Then ${S(t)}$ is strongly but not uniformly convergent on $M$. 
%
\end{enumerate}