%% --
%% -- Updated Notes
%% -- Stand 2025-10-13
%% --
\section*{Updated Notes D-I}
\addcontentsline{toc}{section}{Updated Notes D-I}
%% --
An overview on positive operators on operator algebras can be found in \mycite{zbmath06128372}, but there seems to be no systematic reference for such positive 
$C_{0}$-semigroups on operator algebras.
However, many papers deal with Markov semigroups (see, \eg \mycite{zbmath03762672}) or with so-called E-semigroups (see \mycite{zbmath01949821}).
%% --
\section*{Updated Notes D-II}
\addcontentsline{toc}{section}{Updated Notes D-II}
%% -- 
\begin{enumerate}

\item 
As we have seen in Chapter A-II, Section 3, strongly continuous semigroups on commutative \WA-algebras, that is, on $L^{\infty}$, are already norm-continuous. 
The proof depends heavily on the Grothendieck property (GP) and the Dunford-Pettis property (DPP) of these Banach spaces. 
It was shown by \mycite{zbmath00537321} that every \WA-algebra has (GP) (see also \mycite{zbmath00125315} for an alternative proof).

One can easily see, that \eg $\LH$ with $H$ infinite dimensional does not have (DP) by using the example in D-II.

\item
Let me recall the following: A \WA-algebra is of type I finite iff there exists Hilbert spaces $H_{j}$ of finite dimension and commutative \WA-algebras $M_{j}$ such that 
$M = \oplus_{j} ( \L{H_{j}} \overline{\otimes} M_{j} ) $ (see \mycite[Thm. V.1.27]{zbmath01692441}).

In \mycite{zbmath00097659} it is shown, the predual of every finite type I \WA-algebra has the Dunford-Pettis property.
\mycite{zbmath00125315} have shown that a \WA-algebra $M$ has (DP) iff it is finite of type I such that $ \sup_{j} \dim H_{j} < \infty$.

Thus, as a consequence of A-II Theorem 3.5, in both cases strongly continuous semigroups have a bounded generator.

\item
\mycite{zbmath01495754} has shown that for sequences of completely positive maps, strong and norm convergence to the identity operator are equivalent on \WA-algebras (even on the larger class of \AW-algebras).
Therefore, a strongly continuous $C_{0}$-semigroup of completely positive operators on a \WA-algebra is norm-continuous and thus has a bounded generator.

\item 
In \mycite{zbmath00097715} it is shown that $C(K)$ has the Grothendieck property if and only if each strongly continuous semigroup has a bounded generator. 

If every strongly continuous $C_{0}$-semigroup on a noncommutative \CA-algebra $M$ has a bounded generator, then it is still true that this algebra is a Grothendieck space. 
To prove this, one uses the fact that a \CA-algebra is a Grothendieck space if and only if $c_{0}$ is not a complemented subspace (see 
\mycite[Prop. 3.1.13 and Prop. 4.2.1]{zbmath07458830}).

Indeed, if $M$ is not a Grothendieck-space, we can split $M = c_{0} \oplus F$.
Now define $S(t) = T(t) \oplus Id_{F}$ where $\{T(t)\}$ is strongly but not uniformly continuous on $c_{0}$.
Then $\{S(t)\}$ has an unbounded generator on $M$.

Note that on $\L{H}$ with $H$ an infinite-dimensional Hilbert space, there always exist strongly but not uniformly continuous $C_{0}$-semigroups (see the example in D-II).
However, this \WA-algebra is a Grothendieck space (see again the example at the beginning of D-II).
\end{enumerate}
%%% --
\section*{Updated Notes D-III}
\addcontentsline{toc}{section}{Updated Notes D-III}
%% -- 
\begin{enumerate}
\item 
Since the cone of the self-adjoint, positive elements of a \CA-algebra is a normal cone, one can derive properties such as $s(A) \in \sigma(A)$ or $s(A) = \omega_{0}$ from the theory of semigroups of positive operators on ordered Banach spaces with such cones; see \mycite{zbmath03883065} or \mycite{zbmath03736445}.
But to remain in the category of \CA-algebras, we have summarized this using \CA-algebra methods in a preprint (muss ich noch fertig stellen)

\item 
While \WA-algebras are not stable under the ultraproduct construction, see \mycite[p. 79]{zbmath03640303} or \mycite{zbmath03467832}, their preduals are and can be used for spectral theoretic purposes. 
More on ultraproducts of \WA-algebras can be found in \mycite{zbmath06326930}.

\item 
Another approach to the spectral theory on \WA-algebras is in \mycite{zbmath06031834}, where a Jacobs-de Leeuw-Glicksberg decomposition is constructed.
This leads to a noncommutative version of the Perron-Frobenius theorem for \WA-algebras and is applied to the asymptotics of \WA-dynamical systems.
A similar approach is in \mycite{zbmath06728793}.

\end{enumerate}
%% --
\section*{Updated Notes D-IV}
\addcontentsline{toc}{section}{Updated Notes D-IV}
%% --
As in D-III, results from \mycite{zbmath07964911} on semigroups on ordered Banach spaces with a normal cone and order unit can be applied. 
More precise convergence results on \WA-algebras and their preduals are in \mycite{zbmath02244715}.
Specific investigations focus on \emph{Quantum Markov semigroups and decoherence} (see, \eg, \mycite{zbmath08072228}) or on \emph{Spectral gaps and convergence to equilibrium} (see, \eg \mycite{zbmath06737571}).