%% --
%% -- Updated Notes
%% -- Stand 2025-10-11
%% --
\setlist{$\bullet$, wide, labelindent=.5em,itemsep=.25em}
\section{Updated Notes D-II}
%% -- 
\begin{enumerate}
\item 
As we have seen in Chapter A-II, Section 3, strongly continuous semigroups on commutative \WA-algebras, that is, on $L^{\infty}$, are already norm-continuous. 
The proof depends heavily on the Grothendieck property (GP) and the Dunford-Pettis property (DPP) of these Banach spaces. 
It was shown by \mycite{zbmath00537321} that every \WA-algebra has (GP) (see also \mycite{zbmath00125315} for an alternative proof).
On the other hand, for a \WA-algebra $M$ the following are equivalent.
%% --
\begin{enumerate}[\upshape(a)]
    \item The \WA-algebra $M$ is of Type I finite.
    \item The \WA-algebra $M$ has (DPP).
    \item The predual $M_{*}$ has (DPP).
\end{enumerate}
%% --
(see \mycite{zbmath00125315} and \mycite{zbmath00097659})
Thus, as a consequence of A-II Theorem 3.5, strongly continuous semigroups on Type I finite \WA-algebras have a bounded generator. 
\item
\mycite{zbmath01495754} has shown that strong and norm convergence for sequences of completely positive maps are equivalent on \WA-algebras (even on the larger class of \AW-algebras).
Therefore a strongly continuous $C_{0}$-semigroup of completely positive operators on a \WA-algebra is norm continuous.
\item 
In \mycite{zbmath00097715} it is shown that the equivalence of strong and norm convergence of semigroups on $C(K)$ implies the \CA-algebra $C(K)$ is a Grothendieck space.
This is also true for a noncommutative \CA-algebra $\BA$.
To prove this, one uses the fact that a \CA-algebra is a Grothendieck space if and only if $c_{0}$ is not a complemented subspace (see \mycite[Prop. 3.1.13 \& 4.2.1]{zbmath07458830}).
Indeed, suppose that every strongly convergent $C_{0}$-semigroup is uniformly continuous on $\BA$.
Then $c_{0}$ can't be a complemented subspace of $\BA$.
For if $\BA = c_{0} \oplus F$ for a closed subspace $F$ of $\BA$, then take a strongly convergent but not uniformly convergent semigroup ${T(t)}$ on $c_{0}$ and let $S(t) = T(t) \oplus Id_{F}$ on $\BA$. 
Then ${S(t)}$ is strongly but not uniformly convergent on $\BA$. 
\end{enumerate}

% % **Hauptänderungen:**
% % 1. "non-commutative" → "noncommutative" (amerikanische Schreibweise)
% % 2. "the \CA-algebra" → "the \CA-algebra" (fehlender Artikel ergänzt)
% % 3. "the fact, that" → "the fact that" (Komma entfernt - amerikanischer Stil)

% % Der LaTeX-Code wurde vollständig beibehalten.

% **Vorgenommene Korrekturen:**
% - Zeile 2: Komma nach "that is" hinzugefügt
% - Zeile 5: "is equivalent" → "are equivalent" (Subjekt ist "the following")
% - Zeile 14: "has" → "have" (Subjekt ist "semigroups", Plural)
% - Zeile 17: "$C_{0}$-semigroups" → "$C_{0}$-semigroup" (Singular)
% - Zeile 17: Komma nach "Therefore" hinzugefügt
% - Zeile 20: "a non-commutative" → "for a non-commutative" (fehlende Präposition)
% - Zeile 23: Komma nach "Otherwise" hinzugefügt (Leerzeichen vor $\BA$)
% - Zeile 24: Leerzeichen vor "on" entfernt, Komma nach "$\BA$" hinzugefügt
% - Zeile 24: Komma vor "is strongly" hinzugefügt

\addcontentsline{toc}{section}{Updated Notes: References}